---
title: "Lab5"
author: "Jayson De La O"
format:
  html:
    embed-resources: true
---

```{r}
library(data.table)
library(dtplyr)
library(dplyr)
library(R.utils)
library(leaflet)
```

```{r}
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")
```

```{r}
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]
```

```{r}

# Dealing with NAs and 999999
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]
```

```{r}

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])
```

```{r}

# Dropping NAs
stations <- stations[!is.na(USAF)]
```

```{r}

# Removing duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
```

```{r}
met <- met[,
  .(USAFID, WBAN, year, month, day, hour, min, lat, lon, elev,
    wind.sp, temp, atm.press)]
```

```{r}
metmerge <- merge(
  # Data
  x     = met,      
  y     = stations, 
  # List of variables to match
  by.x  = "USAFID",
  by.y  = "USAF", 
  # Which obs to keep?
  all.x = TRUE,      
  all.y = FALSE
  )
```

QUESTION 1

```{r}

metbys <- metmerge[,.(
  temp     = median(temp,na.rm=TRUE),
  wind.sp  = median(wind.sp,na.rm=TRUE),
  atm.press= median(atm.press,na.rm=TRUE),
  lat      = median(lat, na.rm=TRUE),
  lon      = median(lon, na.rm=TRUE), 
  elev     = median(elev,na.rm=TRUE)
  
), by=.(USAFID)]
metbys



```

Median for wind speed is 2.6, median of temp is 2, and median of atmospheric pressure is 1014.6. The three weather stations that represent the continental US 720277(temp), 720113(wind) , and 720394(atm). The stations did not coincide with the others.

```{r}
stationt <- metbys[which.min(abs(metbys$temp- quantile(metbys$temp,probs=.5,na.rm=TRUE) )) ]

stationt

 stationw<- metbys[which.min(abs(metbys$wind.sp- quantile(metbys$wind.sp,probs=.5,na.rm=TRUE) )) ]
stationw

stationa<- metbys[which.min(abs(metbys$atm.press- quantile(metbys$atm.press,probs=.5,na.rm=TRUE) )) ]
stationa
```

Question 2

```{r}
metstat <- metmerge[,.(
  temp     = median(temp,na.rm=TRUE),
  wind.sp  = median(wind.sp,na.rm=TRUE),
  atm.press= median(atm.press,na.rm=TRUE),
  lat      = median(lat, na.rm=TRUE),
  lon      = median(lon, na.rm=TRUE), 
  elev     = median(elev,na.rm=TRUE)
  
), by=.(USAFID, STATE)]
metstat
```

```{r}

latlong <- metstat %>%
  group_by(STATE) %>%
  summarize(
    avglat=mean(lat),
    avglon=mean(lon),
    avgtemp = median(temp,na.rm=TRUE),
avgwindsp = median(wind.sp,na.rm=TRUE),
avgatmpress = median(atm.press,na.rm=TRUE))
  
latlong
```

States with the Site ID that is the median for that state

```{r}
l <- merge(metstat,latlong, by="STATE")
l


station.dist <-l
station.dist$temp <- station.dist$temp-station.dist$avgtemp
station.dist$wind.sp <- station.dist$wind.sp-station.dist$avgwindsp
station.dist$atm.press <- station.dist$atm.press-station.dist$avgatmpress




station.dist$temp<- station.dist$temp^2
station.dist$wind.sp<- station.dist$wind.sp^2
station.dist$atm.press<- station.dist$atm.press^2



sqr.dist <- station.dist$temp + station.dist$wind.sp + station.dist$atm.press


l$eucl.dist <- sqr.dist
l
 
medstate <- l %>%
  group_by(STATE) %>%

  filter(eucl.dist == min(eucl.dist,na.rm=TRUE)) %>%
  select(STATE,USAFID,lat,lon)
medstate



statemap <- leaflet(medstate) %>%
  addProviderTiles('CartoDB.Positron') %>% 
  # Some circles
  addCircles(
    lat = ~lat, lng=~lon)
                                                
   statemap


```



QUESTION 3
State with the site ID for the station closest to the midpoint
```{r}
l

mid.dist <-l
mid.dist$lon <- abs(mid.dist$lon-mid.dist$avglon)
mid.dist$lat <- abs(mid.dist$lat-mid.dist$avglat)





mid.dist$lon<- mid.dist$lon^2
mid.dist$lat<- mid.dist$lat^2




midsqr.dist <- mid.dist$lon + mid.dist$lat


l$midpoint.dist <- midsqr.dist
l
 
midstate <- l %>%
  group_by(STATE) %>%

  filter(midpoint.dist == min(midpoint.dist,na.rm=TRUE)) %>%
  select(STATE,USAFID,lat,lon)

midstate


midmap <- leaflet(midstate) %>%
  addProviderTiles('CartoDB.Positron') %>% 
  # Some circles
  addCircles(
    lat = ~lat, lng=~lon,color="red")
                                                
   midmap
   
```

```{r}
leaflet() %>%
  addProviderTiles('CartoDB.Positron') %>% 
  # Some circles
  addCircles(data=medstate,
    lat = ~lat, lng=~lon, color="blue") %>%
  addCircles(data=midstate,lat = ~lat, lng=~lon,color="red")
                            
```

Question 4
Summary table that shows num of states, avg temp, avg wind, avg atm.press by avg temp lvl

```{r}
metsummary <- metmerge[,.(
  temp     = mean(temp,na.rm=TRUE),
  wind.sp  = mean(wind.sp,na.rm=TRUE),
  atm.press= mean(atm.press,na.rm=TRUE),
  lat      = mean(lat),
  lon      = mean(lon), 
  elev     = mean(elev,na.rm=TRUE)
  
), by=.( STATE)]
metsummary




metsummary$tc <- ifelse(metsummary$temp  <20,"low",
                       ifelse(metsummary$temp >= 20 & metsummary$temp <25,"Mid",
                              ifelse(metsummary$temp & metsummary$temp >= 25,"High",NA
                      )))
    metsummary                       
```

```{r}
sum <- metsummary %>%
  group_by(tc) %>%
  summarize(
    numstate= sum(table(STATE)),
    lvlavgtemp= mean(temp, na.rm=TRUE),
     lvlavgwind= mean(wind.sp, na.rm=TRUE),
     lvlavgatmpress= mean(atm.press, na.rm=TRUE),
    sum(is.na(metsummary))
  )
sum(is.na(metsummary))
sum
```

